Finiteness in Cubical Type Theory

### 2020-01-25 paper.tex
fixed author names
begining edits

\input{preamble}
\title{Finiteness in Cubical Type Theory}
\author{
  Donnacha Oisín Kidney \inst{1} \and
  Gregory Provan \inst{2}
}
\institute{
  University College Cork \email{o.kidney@cs.ucc.ie} \and
  University College Cork \email{g.provan@cs.ucc.ie}
}
\begin{document}
\maketitle
\begin{abstract}
  We study five different notions of finiteness in Cubical Type Theory and prove
  the relationship between them.
  In particular we show that any totally ordered Kuratowski finite type is
  manifestly Bishop finite.

@@

  We formalise our work in Cubical Agda, where we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying).
  Through this library we demonstrate a number of uses for the computational
  content of the univalence axiom, including searching for and synthesising
  functions.
  \keywords{Agda \and Homotopy Type Theory \and Cubical Type Theory \and
    Dependent Types \and Finiteness \and Topos \and Kuratowski finite}
\end{abstract}

\section{Introduction}
\subsection{Foreword}
In constructive mathematics we are often preoccupied with \emph{why} something
is true.
Take finiteness, for example.
There are a handful of ways to demonstrate some type is finite: we could provide
a surjection from another finite type; we could show that any collection of its
elements larger than some bound contains duplicates; or that any stream of its
elements contain duplicates.

@@

Finally, the other important property of constructive mathematics is that we can
actually \emph{compute} with proofs.
Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise this property
even in the presence of univalence, giving computational content to programs
written in HoTT.
\subsection{Contributions}
In this work we will examine five notions of finiteness in Homotopy Type Theory,
the relationships between them, and their topos-theoretic characterisation.
We also briefly examine a predicate for countable sets, comparing it to the
finiteness predicates.
Our work is formalised in Cubical Agda, where we also develop a library for
proof search based on the finiteness predicates.

>>>>

\input{preamble}
\title{Finiteness in Cubical Type Theory}
\author{
  Donnacha Oisín Kidney\inst{1}\and Gregory Provan \inst{2}
}
\institute{
  University College Cork \email{o.kidney@cs.ucc.ie} \and
  University College Cork \email{g.provan@cs.ucc.ie}
}
\begin{document}
\maketitle
\begin{abstract}
  We study five different notions of finiteness in Cubical Type Theory and prove
  the relationship between them.
  In particular we show that any totally ordered Kuratowski finite type is
  manifestly Bishop finite.

@@

  We formalise our work in Cubical Agda, where we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying).
  Through this library we demonstrate a number of uses for the computational
  content of the univalence axiom, including searching for and synthesising
  functions.
  \keywords{Agda \and Homotopy Type Theory \and Cubical Type Theory \and
    Dependent Types \and Finiteness \and Topos \and Kuratowski finite}
\end{abstract}
\section{Introduction}
\subsection{Foreword}
In constructive mathematics we are often preoccupied with \emph{why} something
is true.
Take finiteness, for example.
There are a handful of ways to demonstrate some type is finite: we could provide
a surjection to it from another finite type; we could show that any collection
of its elements larger than some bound contains duplicates; or we could show
that any stream of its elements contain duplicates.

@@

Finally, for a computer scientist constructive mathematics has one invaluable
feature missing from classical mathematics: computation.
Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise this property
even in the presence of univalence, giving computational content to programs
written in HoTT.
\subsection{Contributions}
In this work we will examine five notions of finiteness in Cubical Type Theory,
the relationships between them, and their topos-theoretic characterisation.
We also briefly examine a predicate for countable sets, comparing it to the
finiteness predicates.
Our work is formalised in Cubical Agda, where we also develop a library for
proof search based on the finiteness predicates.
